### 2020-03-14 Strings2020/strings2/strings2.tex
Add strings2 & update CI

% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

\documentclass[a4paper,10pt]{article}

\newcommand{\hwNumber}{2 [WIP]}

\input{../strings_preamble.tex}
\input{../.modules/basics/biblatex.tex}

%%% ID: sensitive, do NOT publish!
%\InputIfFileExists{../id.tex}{}{}

\newcommand{\oppower}[2]{\mop{{#1}^{#2}}\!}
\newcommand{\sqsinh}{\oppower{\sinh}{2}}
\newcommand{\sqcosh}{\oppower{\cosh}{2}}

\newcommand{\Vir}{\mathbf{V}\mfrak{ir}}
\usepackage{cancel}

\begin{document}
\maketitle
\pagestyle{headings}
\pagenumbering{arabic}
\thispagestyle{empty}

%{
%	\noindent\itshape%
%	本文约定：度规$\eta\sim\pqty{-,+,+,+}$, 指标$\mu,\nu,\dots = 0,1,2,3,\ i,j,\dots = 1,2,3$.
%}
	\begin{enumerate}
	\item \textbf{BRST Quantization of Bosonic String}: 
	\begin{gather}
		S = S^X + S^{bc},\\
		S^X = - \frac{1}{2\pi\alpha'} \int\dd[2]{z}
			\pd X^\mu \pdbar X_\mu,\quad
		S^{bc} = \frac{1}{2\pi} \int\dd[2]{z}
			\pqty\Big{
				b\,\pdbar c
				+ \tilde{b}\,\pd\tilde{c}
			}
	\end{gather}
	This is the gauge fixed action. The corresponding BRST transformation is listed in \textit{Polchinski}; for each of the subsystems, we have its (classical) energy-momentum:
	\begin{equation}
		T^X(z)
		= - \frac{1}{\alpha'}\,
			\normorder{\pd X^\mu \pd X_\mu}\,,\quad
		\tilde{T}^X(\bar{z})
		= - \frac{1}{\alpha'}\,
			\normorder{\pdbar X^\mu \pdbar X_\mu}\,,
		\\[.5ex]
		T^{bc}(z) = \normorder{(\pd b)\,c}
			- 2\,\pd(\normorder{bc}),\quad
		\tilde{T}^{bc}(\bar{z}) = \normorder{
			(\pdbar \tilde{b})\,\tilde{c}
		} - 2\,\pdbar(\normorder{
			\tilde{b}\tilde{c}
		}),
	\end{equation}

	\begin{enumerate}
	\item To get the energy-momentum of $S$, let's visit each of the subsystems respectively; first, BRST transformation of $X$ is given by:
	\begin{equation}
		\var{X^\mu}
		= i\epsilon\,\pqty{
			c\pd + \tilde{c}\pdbar
		} X^\mu
	\end{equation}
	Compared with the conformal transformation\footnote{
		We follow the convention of \textit{Polchinski} unless otherwise stated. 
	}: $
		\var{X^\mu}
		= -\epsilon\,\pqty{
			v\pd
			+ \tilde{v}\pdbar
		} X^\mu
	$, we see that they are in fact identical under the equivalence $
		-\epsilon v \sim i\epsilon c,\ %
		-\epsilon \tilde{v} \sim i\epsilon \tilde{c}
	$, hence we can simply follow the derivation of conformal current and write down $\var{S}^X$'s contribution to the conserved current:
	\begin{equation}
		j^X = c(z)\,T^X(z)
	\end{equation}

	The transformation of $b,c$ is less obvious; for holomorphic current, we need only focus on the holomorphic part of $S^{bc}$; on-shell variation yields:
	\begin{equation}
		0 = \var{S}^{bc}
		= \pqty{\frac{1}{2\pi} \int \dd[2]{z} \pqty{
			- \pdbar c\,\var{b}
			- \pdbar b\,\var{c}
		}}_{\!\!=\mspace{1mu}0}
		+ \frac{1}{2\pi} \int \dd[2]{z}
			\pdbar \pqty{b\var{c}}
		= \frac{1}{2\pi} \int \dd[2]{z}
			\pdbar\epsilon\,\pqty\big{
				-ibc\,\pd c
			} 
	\end{equation}
	Here we've plugged in $
		\var{c} = i\epsilon(z,\bar{z})\,c\pd c
	$, and we have moved $\pdbar\epsilon$ to the beginning of the expression, while respecting the anti-commuting nature of $\epsilon$. With a conventional $i$ coefficent (which agrees with the convention of $j^X$), we have $bc$'s contribution to the conserved current:
	\begin{equation}
		j^{bc} = i\,\pqty\big{
			-ibc\,\pd c
		} = bc\,\pd c
	\end{equation}

	Note that $j^{bc}$ is, in fact, related to the energy-momentum (at least classically):
	\begin{equation}
		\frac{1}{2}\,cT^{bc}
		= \frac{1}{2}\,c\,{(\pd b)\,c}
			- c\,\pd({bc})
		= - c\,\pd({bc})
		= - cb\,\pd c
		= bc\,\pd c = j^{bc}
	\end{equation}
	Hence we have the classical BRST current:
	\begin{equation}
		j(z) = c(z)\,\pqty{
			T^X + \frac{1}{2}\,T^{bc}
		}
	\end{equation}

	For a quantum version, redefine $j(z)$ with normal ordering, and we have:
	\begin{gather}
		T(z)\,j(0)
		\sim T^X(z)\,T^X(0)\,c(0)
			+ T^{bc}(z)\,cT^X(0)
			+ T^{bc}(z)\,
				\normorder{bc\,\pd c}_{\,(0)},
		\label{eq:BRST_Tj_OPE}\\
		\text{where}\quad
		T^X(z)\,T^X(0)\,c(0)
		\sim \pqty{
			\frac{D}{2z^4}
			+ \frac{2}{z^2}\,{T^X(0)}
			+ \frac{1}{z}\,{\pd T^X(0)}
		}\,c(0),
		\label{eq:BRST_Tj_OPE_partI}
	\end{gather}
	Here we've used the fact that $X$ and $b,c$ is de-coupled in the gauge-fixed action, hence their OPE is trivial. Also, we've expanded the first term using $TT$ OPE of the free boson. Additionally, note that $c(z)$ is primary with weight $(-1,0)$, we have:
	\begin{equation}
	\begin{aligned}
		T^{bc}(z)\,cT^X(0)
		&\sim \Bqty{T^{bc}(z)\,c(0)}\,T^X(0) \\
		&\sim \pqty{
			\frac{-1}{z^2}\,c(0)
			+ \frac{1}{z}\,\pd c(0)
		}\,T^X(0),
	\end{aligned}
	\label{eq:BRST_Tj_OPE_partII}
	\end{equation}
	The last term in \eqref{eq:BRST_Tj_OPE} can be brute-forced as follows:
	\begin{equation}
		T^{bc}(z)\,\normorder{bc\,\pd c}_{\,(0)}
		= \pqty\big{
			\normorder{(\pd b)\,c}
			- 2\,\pd(\normorder{bc})
		}_{(z)}\,\normorder{bc\,\pd c}_{\,(0)}\,,
	\end{equation}
	\vspace{-1\baselineskip}
	\begin{align}
		\normorder{(\pd b)\,c}_{\,(z)}\,
		\normorder{bc\,\pd c}_{\,(0)}
		&\sim \normorder{
			\wick{
				(\pd \c b)\,c_{(z)}\,
				b\c c\,\pd c_{(0)}
			}
		} + \normorder{
			\wick{
				(\pd \c b)\,c_{(z)}\,
				bc\,\pd \c c_{(0)}
			}
		} + \normorder{
			\wick{
				(\pd b)\,\c c_{(z)}\,
				\c b c\,\pd c_{(0)}
			}
		} \notag\\
		&\qquad + \normorder{
			\wick{
				(\pd \c1 b)\,\c2 c_{(z)}\,
				\c2 b \c1 c\,\pd c_{(0)}
			}
		} + \normorder{
			\wick{
				(\pd \c1 b)\,\c2 c_{(z)}\,
				\c2 bc \,\pd \c1 c_{(0)}
			}
		} \notag\\[.5ex]
		&\sim \frac{-1}{z^2}\,(+1)\,\normorder{
			c_{(z)}\,b\,\pd c_{(0)}
		} + \frac{-2}{z^3}\,(-1)\,\normorder{
			c_{(z)}\,bc_{(0)}
		} + \frac{1}{z}\,(+1)\,\normorder{
			\pd b_{(z)}\,c\pd c_{(0)}
		} \notag\\
		&\qquad + \frac{-1}{z^2}\cdot\frac{1}{z}\,
			(+1)\,\pd c(0)
		+ \frac{-2}{z^3}\cdot\frac{1}{z}\,
			(-1)\,c(0) \notag\\[.5ex]
		&\sim \frac{-1}{z^2}\,\pqty{
			-j^{bc}(0) + \order*{z^2}
		} + \frac{2}{z^3}\,\pqty{
			z\,j^{bc}(0)
			+ \frac{z^2}{2}\,\normorder{
				bc\,\pd^2 c
			}_{\,(0)}
			+ \order*{z^3}
		} \notag\\[-.3ex]
		&\qquad + \frac{1}{z}\,\pqty\big{
			\normorder{
				(\pd b)\,c\,\pd c
			}_{\,(0)} + \order{z}
		} + \frac{-1}{z^3}\,\pd c(0)
			+ \frac{2}{z^4}\,c(0) \notag\\[.7ex]
		&\sim \frac{4}{2z^4}\,c(0)
			+ \frac{-1}{z^3}\,\pd c(0)
			+ \frac{3}{z^2}\,j^{bc}(0)
			+ \frac{1}{z}\,\normorder{\pqty{
				bc\,\pd^2 c
				+ (\pd b)\,c\,\pd c
			}}_{\,(0)}\,, \notag\\[1ex]
		&\sim \frac{4}{2z^4}\,c(0)
			+ \frac{-1}{z^3}\,\pd c(0)
			+ \frac{3}{z^2}\,j^{bc}(0)
			+ \frac{1}{z}\,\pd j^{bc}(0), \\[3ex]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		THIS IS INSANE!
%		ANOTHER OPE ...
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\normorder{bc}_{\,(z)}\,
		\normorder{bc\,\pd c}_{\,(0)}
		&\sim \normorder{
			\wick{
				\c1 b c_{(z)}\,
				b \c1 c\, \pd c_{(0)}
			}
		} + \normorder{
			\wick{
				\c1 b c_{(z)}\,
				b c\, \pd \c1 c_{(0)}
			}
		} + \normorder{
			\wick{
				b \c1 c_{(z)}\,
				\c1 b c\, \pd c_{(0)}
			}
		} \notag\\
		&\qquad + \normorder{
			\wick{
				\c1 b \c2 c_{(z)}\,
				\c2 b \c1 c\, \pd c_{(0)}
			}
		} + \normorder{
			\wick{
				\c1 b \c2 c_{(z)}\,
				\c2 b c\, \pd \c1 c_{(0)}
			}
		} \notag\\[.5ex]
		&\sim \frac{1}{z}\,(+1)\,\normorder{
			c_{(z)}\,b\,\pd c_{(0)}
		} + \frac{1}{z^2}\,(-1)\,\normorder{
			c_{(z)}\,bc_{(0)}
		} + \frac{1}{z}\,(+1)\,\normorder{
			b_{(z)}\,c\pd c_{(0)}
		} \notag\\
		&\qquad + \frac{1}{z}\cdot\frac{1}{z}\,
			(+1)\,\pd c(0)
		+ \frac{1}{z^2}\cdot\frac{1}{z}\,
			(-1)\,c(0) \notag\\[1ex]
		&\sim \frac{1}{z}\,\pqty\big{
			-j^{bc}(0)
		} + \frac{-1}{z^2}\,\pqty\big{
			z\,j^{bc}(0)
		} + \frac{1}{z}\,\pqty\big{
			j^{bc}(0)
		} + \frac{1}{z^2}\,\pd c(0)
		+ \frac{-1}{z^3}\,c(0) \notag\\[1ex]
		&\sim \frac{-1}{z^3}\,c(0)
			+ \frac{1}{z^2}\,\pd c(0)
			+ \frac{-1}{z}\,j^{bc}(0), \\[2ex]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		DERIVATIVE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\pd (\normorder{bc})_{(z)}\,
		\normorder{bc\,\pd c}_{\,(0)}
		&\sim \frac{6}{2z^4}\,c(0)
			+ \frac{-2}{z^3}\,\pd c(0)
			+ \frac{1}{z^2}\,j^{bc}(0),
	\end{align}
	\begin{gather}
		T^{bc}(z)\,\normorder{bc\,\pd c}_{\,(0)}
		\sim \frac{-8}{2z^4}\,c(0)
			+ \frac{3}{z^3}\,\pd c(0)
			+ \frac{1}{z^2}\,j^{bc}(0)
			+ \frac{1}{z}\,\pd j^{bc}(0),
		\label{eq:BRST_Tj_OPE_partIII}\\[2ex]
		T(z)\,j(0)
		\sim \eqref{eq:BRST_Tj_OPE_partI}
			+ \eqref{eq:BRST_Tj_OPE_partII}
			+ \eqref{eq:BRST_Tj_OPE_partIII}
		\sim \frac{D-8}{2z^4}\,c(0)
			+ \frac{3}{z^3}\,\pd c(0)
			+ \frac{1}{z^2}\,j(0)
			+ \frac{1}{z}\,\pd j(0),
	\end{gather}

	We see that $j(z)$ defined with na\"ive normal ordering is \textit{almost} but \textit{not quite} a primary. It differs from primary OPE at $\order{\frac{1}{z^4}}$ and $\order{\frac{1}{z^3}}$. However, it is possible to make it into a primary by adding extra terms that do not interfere with current conservation $\pdbar j = 0$. To cancel the $\frac{3}{z^3}\,\pd c(0)$ term, notice that $
		b(z)\,\pd^2 c(0) \sim \frac{2}{z^3}
	$, therefore it may be helpful to look at:
	\begin{equation}
	\begin{aligned}
		T(z)\,\pd^2 c(0)
		\sim T^{bc}(z)\,\pd^2 c(0)
		&\sim \pd^2_w \pqty{
			T^{bc}(z)\,c(w)
		}_{w\to 0} \\
		&\sim \pd^2_w \pqty{
			\frac{-1}{(z-w)^2}\,c(w)
			+ \frac{1}{z-w}\,\pd c(w)
		}_{\!w\to 0} \\[.5ex]
		&\sim \frac{-12}{2z^4}\,c(0)
			+ \frac{-2}{z^3}\,\pd c(0)
			+ \frac{1}{z^2}\,\pd^2 c(0)
			+ \frac{1}{z}\,\pd^3 c(0),
	\end{aligned}
	\end{equation}
	Again we've used $Tc$ OPE of the primary $c(w)$. We see that indeed, the $\frac{1}{z^3}\,\pd c(0)$ term can be canceled by shifting $j(z)$:
	\begin{gather}
		j(z)\ \longmapsto\ j(z)
			+ \frac{3}{2}\,\pd^2 c(z),\quad
		j(z) = j^X + j^{bc}
			+ \frac{3}{2}\,\pd^2 c,\\[1ex]
		T(z)\,j(0)
		\sim \frac{D-26}{2z^4}\,c(0)
			+ \frac{1}{z^2}\,j(0)
			+ \frac{1}{z}\,\pd j(0),
	\end{gather}

	We see that $j(z)$ defined in this way is a primary of weight $(1,0)$ in $D = 26$. This is the quantum BRST current. 
	\qedfull

%	note that $b,c$ are primaries with weight $(2,0)$ and $(-1,0)$, hence $Tb,Tc$ OPEs take the standard form of a $T\cdot(\text{primary})$ OPE. We have:
%	\begin{equation}
%	\begin{aligned}
%		T^{bc}(z)\,\normorder{bc\,\pd c}_{\,(0)}
%		&\sim \pqty{
%				\frac{2}{z^2}\,
%					\normorder{bc\,\pd c}
%				+ \frac{1}{z}\,\normorder{
%					(\pd b)\,c\pd c
%				}
%			} \\
%		&\phantom{\sim \quad}
%			+ \pqty{
%				\frac{-1}{z^2}\,
%					\normorder{bc\,\pd c}
%				+ \frac{1}{z}\,\normorder{
%					b\,(\pd c)(\pd c)
%				}
%			} \\
%		&\phantom{\sim \quad}
%			+ \pqty{
%				\frac{-1}{z^2}\,
%					\normorder{bc\,\pd c}
%				+ \frac{1}{z}\,\normorder{
%					bc\,\pd^2 c
%				}
%			} \\
%		&\sim \frac{1}{z}\,\pd\,
%			\normorder{bc\,\pd c}
%		\sim \frac{1}{z}\,\pd\,\pqty{
%			j - cT^X
%		}_{(0)} \\
%		&\sim \frac{1}{z}\,\pd j(0)
%		- \frac{1}{z}\,c(0)\,\pd T^X(0)
%		- \frac{1}{z}\,T^X(0)\,\pd c(0),
%	\end{aligned}\\
%		T(z)\,j(0)
%		\sim \frac{D}{2z^4}\,c(0)
%			+ \frac{2}{z^2}\,T^X(0)\,c(0)
%			+ \frac{1}{z}\,\pqty{
%				\pd j(0)
%				- T^X(0)\,\pd c(0)
%			}
%	\end{equation}

	\item

	\end{enumerate}

	\item \textbf{Linear Dilaton CFT:}

	\clearpage

	\quad
	\vfill\pagebreak

	\item \textbf{Spinning Closed String in AdS Space}:

	For a classical spinning string, we have Nambu--Goto action:
	\begin{equation}
		S_{NG}
		= -T \int \dd{\tau} \dd{\sigma}
			\sqrt{-\det \gamma_{ab}},\quad
		\gamma_{ab} = G_{\mu\nu}
			\pdd{a}X^\mu \pdd{b}X^\nu
	\end{equation}
	Here $G_{\mu\nu}$ is the spacetime metric. $\gamma_{ab}$ can be treated as the induced metric on the worldsheet.

	In AdS space we have:
	\begin{equation}
		ds^2
		= R^2\pqty{
			- \sqcosh\rho \dd{t}^2
			+ \dd{\rho}^2
			+ \sqsinh\rho \dd{\Omega}^2
		}
	\end{equation}
	Where $\dd{\Omega}^2$ is the metric of a unit $(d-2)$--sphere $S^{d-2}$. For convenience let's define unit $S^{d-2}$ metric $G^1_{ij}$, and raise or lower the $i,j,\cdots$ indices using $G^1_{ij}$ instead of $G_{ij}$, i.e.,
	\begin{equation}
		G^1_{ij} = G_{ij} /\,\pqty{
			R^2 \sqsinh\rho
		},\quad
		i,j = 2,\cdots,d-1
	\end{equation}

	Furthermore, we consider the special case that the closed string is \textit{folded}, like a rubber band stretched along a line; in this case we can choose the worldsheet parameter $(\tau,\sigma) = (t,\rho)$ while $\Omega = \Omega(t,\rho) = \Omega(\tau,\sigma)$, which leads to the following decomposition:
	\begin{align}
		\pdd{a} X^\mu
		&= \delta^\mu_a + \delta^\mu_i\,
			\pdd{a}\Omega^i,\quad
			a = 0,1,\quad
			i = 2,\cdots,d-1,\\[.5ex]
		\gamma_{ab}
		&= G_{\mu\nu}\,
			\pdd{a}X^\mu \pdd{b}X^\nu
		\notag\\
		&= G_{ab} + G_{ij}\,
			\pdd{a}\Omega^i\,\pdd{b}\Omega^j
		\notag\\
		&= G_{ab} + R^2\sqsinh\rho\,G^1_{ij}\,
			\pdd{a}\Omega^i\,\pdd{b}\Omega^j
		\notag\\
		&= R^2\Bqty{\mat{
			-\sqcosh\rho & \\
			& 1 \\
		} + \sqsinh\rho\,\mat{
			(\pdd{a}\Omega)^2 &
			\pdd{a}\Omega\cdot\pdd{b}\Omega \\
			\pdd{b}\Omega\cdot\pdd{a}\Omega &
			(\pdd{b}\Omega)^2 \\
		}}
%		\notag\\
%		&= R^2\,\sqsinh\rho\,\Bqty{\mat{
%			-\oppower{\coth}{2}\rho & \\
%			& \oppower{\csch}{2}\rho \\
%		} + \,\mat{
%			(\pdd{a}\Omega)^2 &
%			\pdd{a}\Omega\cdot\pdd{b}\Omega \\
%			\pdd{b}\Omega\cdot\pdd{a}\Omega &
%			(\pdd{b}\Omega)^2 \\
%		}}
	\end{align}
	Here $
		\pdd{a}\Omega\cdot\pdd{b}\Omega
		\equiv \pdd{a}\Omega^i\,\pdd{b}\Omega_i
		\equiv G^1_{ij}\,
			\pdd{a}\Omega^i\,\pdd{b}\Omega^j
	$, and we have:
	\begin{equation}
	\begin{aligned}
		\det \gamma_{ab}
		&= (R^2)^2\,\Big\{
			\oppower{\sinh}{4}\rho\,
			\det\,(\pdd{a}\Omega^i\pdd{b}\Omega_i)
		\\
		&\qquad\qquad\qquad
			+ \oppower{\sinh}{2}\rho\,\pqty{
				(\pdd{a}\Omega)^2
				- (\pdd{b}\Omega)^2
				\oppower{\cosh}{2}\rho
			}
		\\
		&\qquad\qquad\qquad
			- \oppower{\cosh}{2}\rho\,
		\Big\},\\[.5ex]
		\sqrt{-\det \gamma_{ab}}
		&= R^2\,\Big\{
			\oppower{\cosh}{2}\rho
			- \oppower{\sinh}{2}\rho\,\pqty{
				(\pdd{a}\Omega)^2
				- (\pdd{b}\Omega)^2
				\oppower{\cosh}{2}\rho
			}
		\\[-.5ex]
		&\qquad\qquad\quad
			- \oppower{\sinh}{4}\rho\,
			\det\,(\pdd{a}\Omega^i\pdd{b}\Omega_i)
		\Big\}^{\!1/2}
	\end{aligned}
	\end{equation}

	Mark the end points of the string with $\rho = r(t)$, then the total length of such closed folded string is $\ell = 4r$. We then have:
	\begin{equation}
		S = -4TR^2 \int\dd{t} \int_0^r \dd{\rho}
		\sqrt{
			\oppower{\cosh}{2}\rho
			- \oppower{\sinh}{2}\rho\,\pqty{
				(\pdd{a}\Omega)^2
				- (\pdd{b}\Omega)^2
				\oppower{\cosh}{2}\rho
			}
			- \oppower{\sinh}{4}\rho\,
			\det\,(\pdd{a}\Omega^i\pdd{b}\Omega_i)
		}
	\end{equation}
\pagebreak[3]

%	Variation of metric determinant $
%		\gamma \equiv \det\gamma_{ab}
%	$ is give by the following formula:
%	\begin{equation}
%		\var{\sqrt{-\gamma}}
%		= \frac{1}{2\sqrt{-\gamma}}
%			\var{(-\gamma)}
%		= \frac{1}{2\sqrt{-\gamma}}\,
%			(-\gamma)
%			\var{\ln\gamma}
%		= \frac{1}{2} \sqrt{-\gamma}\,
%			\gamma^{ab} \var{\gamma_{ab}}
%	\end{equation}

	Further simplification comes from the fact that, due to rotational symmetry, the string's motion can be restricted in a plane where its position is characterized by some angle $\theta = \Omega^{i_0}\in \{\Omega^i\}_i$. In this case other angle parameters $\Omega^i|_{i\ne i_0} = 0$, and the action is further reduced to:
	\begin{gather}
		S = -4TR^2 \int\dd{t} \int_0^r \dd{\rho}
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \oppower{\sinh}{2}\rho\,\pqty{
					(\pdd{a}\theta)^2
					- (\pdd{b}\theta)^2
					\oppower{\cosh}{2}\rho
				}
			}
		= \int\dd{t} \int_0^r \dd{\rho} \mcal{L},\\
		\mcal{L} = -4TR^2
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			},\quad
			\omega = \pdd{t}\theta,\,
			\pdd{\rho}\theta = 0
	\end{gather}
	We consider the special solution $\theta = \omega t$, while in general the endpoint $r = r(t)$ could be dynamical; variation of the action w.r.t.\ $r(t)$ gives\footnote{
		The above reasoning is confirmed in e.g.\ \arxiv{hep-th/0204051}.
	}:
	\begin{gather}
		0 = \var{S} = -4TR^2 \int\dd{t}
			\int\limits_r^{r+\var{r}} \dd{\rho}
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			}
		= -4TR^2 \int\dd{t}
			\sqrt{
				\oppower{\cosh}{2}r
				- \omega^2\oppower{\sinh}{2}r
			} \var{r},\\
		\omega^2 = \frac{\sqcosh r}{\sqsinh r}
		= \oppower{\coth}{2} r
		\label{eq:string_endpt_constraint}
	\end{gather}

	Note that if $\omega$ is constant, then $r$ must be fixed by \eqref{eq:string_endpt_constraint}. 
	Taking $\theta$ as the only dynamical variable, it is then straight-forward to write the energy $E$ and angular momentum $J$ for such folded closed string:
	\begin{gather}
		\omega = \dot{\theta},\quad
		\Pi = \pdv{\mcal{L}}{\omega}
		= 4TR^2\,\frac{\omega\sqsinh\rho}{
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			}
		},\\
		J = \int_0^r \dd{\rho} \Pi
		= 4TR^2 \int_0^r \dd{\rho}
		\frac{\omega\sqsinh\rho}{
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			}
		},\\
		E = \int_0^r \dd{\rho} \pqty{
			\Pi\omega - \mcal{L}
		}
		= 4TR^2 \int_0^r \dd{\rho}
		\frac{\sqcosh\rho}{
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			}
		},
	\end{gather}

	In the large string limit, $r\to\infty,\ \omega = \coth r\to 1$. Expand in terms of $\epsilon = \omega - 1 > 0$, we find that $
		r = \frac{1}{2} \ln\pqty{
			1 + \frac{2}{\epsilon}
		}
		\sim \frac{1}{2} \ln\frac{2}{\epsilon}
	$, or alternatively, $e^{2r}\cdot\epsilon\sim 2$. With some help from Mathematica\texttrademark, we get:
	\begin{equation}
	\begin{aligned}
		E - J
		&= 4TR^2 \int_0^r \dd{\rho}
		\frac{\sqcosh\rho - \omega\sqsinh\rho}{
			\sqrt{
				\oppower{\cosh}{2}\rho
				- \omega^2\oppower{\sinh}{2}\rho
			}
		}
%		\\[.5ex]
%		&
		= 4TR^2 \int_0^r \dd{\rho} \pqty\Big{
			1 + \frac{\epsilon^2}{8}\sinh^2(2\rho)
			+ \order{\epsilon^3}
		} \\[1ex]
		&= 4TR^2 \pqty{
			r \pqty\Big{
				1 - \frac{\epsilon^2}{16}
				+ \order{\epsilon^3}
			} + \order{1}
		}
%		\\[.5ex]
%		&
		= \pqty{2TR^2 \ln\frac{2}{\epsilon}}
		\pqty\Big{
			1 - \frac{\epsilon^2}{16}
			+ \order{\epsilon^3}
		} \\[1ex]
		&\sim 2TR^2
			\pqty{\ln\frac{2}{\epsilon}}
	\end{aligned}
	\end{equation}
	Similarly, $
		J \sim 4TR^2 \mathlarger{\int_0^r}
			\dd{\rho} \sqsinh \rho
		\sim TR^2 \pqty{\frac{2}{\epsilon}}
	$, this gives:
	\begin{equation}
		E - J
		\sim 2TR^2 \ln \frac{J}{TR^2}
	\end{equation}
	\qedfull
\pagebreak[4]

	\item \textbf{Special Conformal Transformations:}
	\begin{equation}
		x^\mu
		\xmapsto{\ K(a)\ }
		\tilde{x}^\mu
		= \frac{x^\mu + x^2 a^\mu}{
			1 + 2a\cdot x + a^2 x^2
		}
	\end{equation}

	\begin{enumerate}
	\item Under special conformal transformation $K(a)$, metric $\delta_{\mu\nu}\mapsto g_{\mu\nu}$ while:
	\begin{equation}
		g_{\alpha\beta}
			\dd{\tilde{x}^\alpha}
			\dd{\tilde{x}^\beta}
		= \delta_{\mu\nu}
			\dd{x^\mu}
			\dd{x^\nu},\quad
		g_{\alpha\beta}
		= \delta_{\mu\nu}
			\pdv{x^\mu}{\tilde{x}^\alpha}
			\pdv{x^\nu}{\tilde{x}^\beta}
		\label{eq:metric_transform}
	\end{equation}
	To calculate this we have to know the inverse transformation $x = K^{-1}(a)\,\tilde{x}$. First, notice the following decomposition\footnote{
		See \textit{Di Francesco et al}, and also \https{github.com/davidsd/ph229}. 
	} of $K(a)$:
	\begin{equation}
		\tilde{x}^\mu
		= \frac{\frac{x^\mu}{x^2} + a^\mu}{
			\frac{1}{x^2}
			+ \frac{2a\cdot x}{x^2} + a^2
		}
		= \frac{\frac{x^\mu}{x^2} + a^\mu}{
			\abs{\frac{x^\mu}{x^2} + a^\mu}^2
		},\\[1ex]
		\text{i.e.}\quad
		K(a)\colon\ %
			x^\mu
			\ \xmapsto{\ I\ }\ %
			\frac{x^\mu}{x^2}
			\ \xmapsto{\ T(a)\ }\ %
			y^\mu = \frac{x^\mu}{x^2} + a^\mu
			\ \xmapsto{\ I\ }\ %
			\tilde{x}^\mu
			= \frac{y^\mu}{y^2},
		\label{eq:special_conformal_decomp} \\[.5ex]
		\text{i.e.}\quad
		\frac{\tilde{x}^\mu}{\tilde{x}^2}
			= \flatfrac{
				\frac{y^\mu}{y^2}
			}{
				\frac{1}{y^2}
			}
			= y^\mu = \frac{x^\mu}{x^2} + a^\mu
		\label{eq:special_conformal_simplified}
	\end{equation}

	From \eqref{eq:special_conformal_simplified}, we see that the transformation parameter $a^\mu$ composes linearly: $K(b)\,K(a) = K(a+b)$, therefore $K^{-1}(a) = K(-a)$, and we have:
	\begin{equation}
		x^\mu = K(-a)\,\tilde{x}^\mu
		= \frac{\tilde{x}^\mu - \tilde{x}^2 a^\mu}{
			1 - 2a\cdot \tilde{x} + a^2 \tilde{x}^2
		}
		= \frac{\tilde{y}^\mu}{y^2},\\[1.5ex]
	\begin{aligned}
		\pdv{x^\mu}{\tilde{x}^\alpha}
		&= \pdv{x^\mu}{\tilde{y}^\sigma}
			\pdv{\tilde{y}^\sigma}{\tilde{x}^\alpha}
		= \pqty{
			\pdv{\tilde{y}^\sigma}
			\frac{\tilde{y}^\mu}{\tilde{y}^2}
		} \pdv{\tilde{x}^\alpha} \pqty{
			\frac{\tilde{x}^\sigma}{\tilde{x}^2}
			- a^\sigma\!
		}
		= \pqty{
			\pdv{\tilde{y}^\sigma}
			\frac{\tilde{y}^\mu}{\tilde{y}^2}
		} \pqty{
			\pdv{\tilde{x}^\alpha}
			\frac{\tilde{x}^\sigma}{\tilde{x}^2}
		} \\[1ex]
		&= \pqty{
			\tilde{y}^2\delta^\mu_\sigma
			- 2\tilde{y}^\mu \tilde{y}_\sigma
		} \pqty{
			\tilde{x}^2\delta^\sigma_\alpha
			- 2\tilde{x}^\sigma \tilde{x}_\alpha
		} \Big/ \pqty{\tilde{y}^4 \tilde{x}^4},
	\end{aligned}\\[1ex]
	\begin{aligned}
		g_{\alpha\beta}
		&\xlongequal{\eqref{eq:metric_transform}}
		\delta_{\mu\nu} \pqty{
			\tilde{y}^2\delta^\mu_\sigma
			- 2\tilde{y}^\mu \tilde{y}_\sigma
		} \pqty{
			\tilde{x}^2\delta^\sigma_\alpha
			- 2\tilde{x}^\sigma \tilde{x}_\alpha
		} \pqty{
			\tilde{y}^2\delta^\nu_\rho
			- 2\tilde{y}^\nu \tilde{y}_\rho
		} \pqty\big{
			\tilde{x}^2\delta^\rho_\beta
			- 2\tilde{x}^\rho \tilde{x}_\beta
		} \Big/ \pqty{\tilde{y}^8 \tilde{x}^8}\\
		&\xlongequal{\sum_{\mu,\nu}}
		\tilde{y}^{-4} \delta_{\sigma\rho} \pqty{
			\tilde{x}^2\delta^\sigma_\alpha
			- 2\tilde{x}^\sigma \tilde{x}_\alpha
		} \pqty\big{
			\tilde{x}^2\delta^\rho_\beta
			- 2\tilde{x}^\rho \tilde{x}_\beta
		} \Big/ \tilde{x}^8 \\
		&\xlongequal{\sum_{\sigma,\rho}}
			\tilde{y}^{-4} \tilde{x}^{-4}
			\delta_{\alpha\beta}
	\end{aligned}
	\end{equation}
	We see that $g_{\alpha\beta} = f(x)\,\delta_{\alpha\beta}$, with coefficient:
	\begin{equation}
		f(x) = \tilde{y}^{-4} \tilde{x}^{-4}
		\xlongequal{%
			\eqref{eq:special_conformal_decomp}%
		} \frac{x^4}{\tilde{x}^4}
		\xlongequal{%
			\eqref{eq:special_conformal_simplified}%
		} \pqty{
			1 + 2a\cdot x + a^2 x^2
		}^2
	\end{equation}
	\qed[(a)]
	\vspace*{-1\baselineskip}

%	Note that the transformation depends smoothly on parameter $a$, therefore we can consider infinitesimal transformation generator:
%	\begin{equation}
%		k_\nu
%		= k\id{^\mu_\nu}\pdd{\mu}
%		= \pdv{\tilde{x}^\mu}{a^\nu}
%			\bigg|_{a=0} \!\pdd{\mu}
%		= \pqty{x^2\delta^\mu_\nu
%			- 2x^\mu x_\nu}\pdd{\mu}
%	\end{equation}

	\item In 2D with $z = x^1 + ix^2,\ x^\mu\sim (z,\bar{z})$, we see from \eqref{eq:special_conformal_simplified} that:
	\begin{equation}
		\frac{x^\mu}{x^2}
		\sim \frac{z}{\abs{z}^2}
		= \frac{1}{\bar{z}}
		\ \longmapsto\ %
		\frac{1}{\bar{z}} + a,\quad
		\text{i.e.}\quad
		z\ \longmapsto\ 
		w = \frac{1}{\frac{1}{z} + \bar{a}}
		= \frac{z}{1 + z\bar{a}}
	\end{equation}
	Expand in the $\bar{a}\to 0$ limit, we find that $
		w = z\,\pqty{
			1 - z\bar{a} - \cdots
		} \sim z - z^2 \bar{a}
	$, i.e.\ it is generated by:
	\begin{equation}
		K_{\bar{z}}
		= -z^2 \pdd{z}
		= -z^2 \pd,\quad
		\pd \equiv \pdd{z}
	\end{equation}
\pagebreak[3]

	Note that when considering non-holomorphic functions, we have to consider $(z,\bar{z})$ as \textit{two} independent variables; hence the anti-holomorphic transformation $
		\bar{z}\mapsto\bar{w}
		= \frac{\bar{z}}{1 + \bar{z} a}
		\sim \bar{z} - \bar{z}^2 a
	$ provides another degree of freedom, namely:
	\begin{gather}
		K_\mu\ \sim\ \pqty{
			K_{\bar{z}} = -z^2 \pd,\ %
			K_{{z}} = -\bar{z}^2 \pdbar
		},\\
		\pd \equiv \pdd{z},\ %
		\pdbar \equiv \pdd{\bar{z}}\notag
	\end{gather}
	Similarly, for translation $z\mapsto z+a$ and its conjugate, we have $
		P_\mu\sim
		\pqty{P_z = \pd, P_{\bar{z}} = \pdbar}
	$. However, dilation and rotation are both encoded in a complex rescaling $
		z\mapsto \lambda z,\ %
		\lambda = re^{i\theta}\in\mbb{C}
	$; we have:
	\begin{equation}
		z\mapsto \lambda z,\quad
		\lambda = re^{i\theta}\in\mbb{C},\quad
	\begin{aligned}
		\var{r} \ \longleftrightarrow\ %
		D &= z\,\pd + \bar{z}\,\pdbar,\\[-.5ex]
		\var{\theta} \ \longleftrightarrow\ %
		M &= i\,\pqty{z\,\pd - \bar{z}\,\pdbar},
	\end{aligned}
	\end{equation}

	In summary, we have $
		\mop{span}_\mbb{R}\Bqty{
			P_\mu, K_\mu, D, M
		} = \mfrak{so}(3,1)
	$ generating the ``global'' transformation subgroup of the 2D conformal group; here, the $\mfrak{so}(3,1)$ boost is a linear combination\footnote{
		See e.g.\ \https{github.com/davidsd/ph229}. 
	} of $P_\mu$ and $K_\mu$. More specifically, in 2D any holomorphic or anti-holomorphic function gives a conformal transformation, hence the (classical) 2D conformal group is generated by:
	\begin{equation}
		\ell_m = z^{m+1} \pd,\quad
		\bar{\ell}_m = \bar{z}^{m+1} \pdbar,\quad
		m\in\mbb{Z}
	\end{equation}
	i.e.\ the \textit{Witt algebra} (or Virasoro algbera $\Vir_c$ with $c = 0$). It is clear that a (complexified) $\mfrak{so}(3,1)$ lives inside $\Vir_c$, i.e.,
	\begin{equation}
	\begin{aligned}
		\mfrak{so}(3,1)^\mbb{C}
		&= \mop{span}_\mbb{C}\Bqty{
			P_\mu, K_\mu, D, M
		} \\
		&= \mop{span}_\mbb{C}\Bqty{
			\ell_m,\bar{\ell}_m
			\,\big|\, m = 0,\pm 1
		}
		= \mfrak{sl}(2,\mbb{R})^\mbb{C}
			\oplus_\mbb{C}
			\mfrak{sl}(2,\mbb{R})^\mbb{C}
		\subset \Vir_{c}
	\end{aligned}
	\end{equation}
	\end{enumerate}
	\vspace{-1.2\baselineskip}

	\qedfull

	\item \textbf{$bc$ CFT}:
	\begin{equation}
		S = \frac{1}{2\pi}
			\int \dd[2]{z} b\,\pdbar c
	\end{equation}

	Stress tensor of a theory can be obtained via variation over the metric, or equivalently, over the fields $\phi^i$ with $\var{\phi}$ induced by some \textit{local} spacetime translation $
		x^\mu\mapsto x^\mu + \var{x}^\mu,\ %
		\var{x}^\mu = \epsilon(x)\,a^\mu
	$. Here $\epsilon(x)$ is any compactly supported bump function, centered around some point $x_0$.

	In 2D, we have $\mu = z,\bar{z}$; for $\phi(z,\bar{z})$ with conformal weight $(h,\bar{h})$, consider $
		z\mapsto z',\,
		\bar{z}\mapsto \bar{z}'
	$. For convenience, let's first consider a generic variation $\var{z} = \epsilon(z,\bar{z})$ before restricting to spacetime translation; we have:
	\begin{equation}
		\phi'(z',\bar{z}')
		= \pqty\Big{\dv{z'}{z}}^{\!\! -h}
			\pqty\Big{\dv{\bar{z}'}{\bar{z}}}^{
				\!\! -\bar{h}
			}
			\phi(z,\bar{z}),\\[1ex]
		\tilde{\delta}\phi
		= \pqty{
			- h\,\pd\epsilon
			- \bar{h}\,\pdbar\bar{\epsilon}\,
		}\,\phi,\\
		\var{\phi} 
		= \tilde{\delta}\phi
			-\pdv{\phi}{x^\mu} \var{x^\mu}
		= \pqty{
			- h\,\pd\epsilon
			- \bar{h}\,\pdbar\bar{\epsilon}\,
		}\,\phi
			- \epsilon\,\pd\phi
			- \bar{\epsilon}\,\pdbar\phi,
		\label{eq:var_field_compact_supp}
	\end{equation}
	Here we use $\tilde{\delta}\phi$ to denote the ``internal'' variation related to the conformal weights.

	Note that $\phi = b,c$ are anti-commuting Grassmann numbers, variation of the action gives:
	\begin{equation}
	\begin{aligned}
		\var{S}[b,c,\var{b},\var{c}]
		&= \frac{1}{2\pi} \int \dd[2]{z} \pqty{
			\var{b}\,\pdbar c
			+ b\,\pdbar \var{c}
		} \\
		&= \frac{1}{2\pi} \int \dd[2]{z} \pqty{
			- \pdbar c\,\var{b}
			- \pdbar b\,\var{c}
		} + \frac{1}{2\pi} \int \dd[2]{z}
			\pdbar \pqty{b\var{c}}
	\end{aligned}
	\label{eq:bc_action_var}
	\end{equation}
	For \textit{unknown} $b,c$ and {arbitary} $\var{b},\var{c}$, the second term is reduced to a boundary term at infinity and can be dropped; imposing $\var{S} = 0$ gives the equation of motion (EOM): $\pdbar b = \pdbar c = 0$.

	On the other hand, for \textit{on-shell} $b,c$ and compactly supported $\varphi = \var{b},\var{c}$ given in \eqref{eq:var_field_compact_supp}, the first term in \eqref{eq:bc_action_var} vanishes while $\var{S}_0 = 0$ still holds; this gives:
	\begin{equation}
	\begin{aligned}
		0 = \var{S}_0
		&= \frac{1}{2\pi} \int \dd[2]{z}
			\pdbar \pqty{b\var{c}}
		= \frac{1}{2\pi} \int \dd[2]{z}
			\pdbar\,\pqty\big{
				- (1-\lambda)\,bc\,\pd\epsilon
				- b\,\pd c\,\epsilon
			} \\
		&= \frac{1}{2\pi} \int \dd[2]{z}
			\pqty\big{
				- (1-\lambda)\,bc\,
					\pdbar\pd\epsilon
				- b\,\pd c\,\pdbar\epsilon
			}
	\end{aligned}
	\end{equation}
	Here we've distributed the $\pdbar$ operator and dropped all terms that vanish automatically by EOM. Next we shall collect the $\pd\epsilon,\pdbar\epsilon$ terms; integrating by parts on the first integrand gives:
	\begin{equation}
	\begin{aligned}
		0 = \var{S}_0
		&= \frac{1}{2\pi} \int \dd[2]{z}
			\pqty\big{
				(1-\lambda)\,\pd(bc)\,
				- b\,\pd c
			}\,\pdbar\epsilon \\
		&= \frac{1}{2\pi} \int \dd[2]{z}
			\pqty\big{
				(\pd b)\,c
				- \lambda\,\pd(bc)
			}\,\pdbar\epsilon \\
		&= - \frac{1}{2\pi} \int \dd[2]{z}
			\epsilon(z,\bar{z})\,\pd_{\bar{z}}
			\pqty\big{
				(\pd b)\,c
				- \lambda\,\pd(bc)
			}
	\end{aligned}
	\label{eq:conserved_current}
	\end{equation}

	Notice that we have obtained a conserved current using a generic $
		\var{z}
		= \epsilon(z,\bar{z}),
		\var{\bar{z}}
		= \bar{\epsilon}(z,\bar{z})
	$; by setting $\epsilon = \epsilon(z)$, we get \textit{a} energy momentum tensor\footnote{
		Note that the energy momentum tensor obtained in this way is generally \textit{not} unique: it can be off by a boundary term; see Luboš' comment at \https{physics.stackexchange.com/a/96100}, also \arxiv{1601.03616}. However, it is possible to fix this redundancy by considering $Tb$ OPE and match its conformal dimension. I would like to thank \textit{林般} for pointing this out. 
	}:
	\begin{equation}
		T(z)
		= \normorder{(\pd b)\,c}
			- \lambda\,\pd(\normorder{bc})
	\end{equation}
	Normal ordering is added manually to remove singular terms.

	To compute $TT$ OPE, we need the OPE of $b(z)\,c(0)$; this is obtained by examining the following path integral, which is zero since the integrand is a total functional derivative:
	\begin{equation}
		0 = \int \mscr{D}b\,\mscr{D}c\,
			\fdv{\phi}\pqty{e^{-S}\,\psi}
	\end{equation}
	Taking $\phi,\psi=b,c$, this generates operator equations such as $
		\pdbar b(z)\,c(0)
		= 2\pi\delta^2(z,\bar{z})
	$. Note that $
		\pdbar(\frac{1}{z})
		= 2\pi\delta^2(z,\bar{z})
	$, which gives:
	\begin{equation}
		b(z)\,c(0) \sim c(z)\,b(0)
		\sim \frac{1}{z},\quad
		b(z)\,b(0) \sim 0
		\sim c(z)\,c(0)
	\end{equation}

	With the $bc$ OPE in hand, the $TT$ OPE is computed directly with brute force, by repeatedly applying Wick's theorem. This gives:
	\begin{equation}
		T(z)\,T(0)
		\sim \frac{
			-6\lambda^2 + 6\lambda - 1
		}{z^4} + \cdots
	\end{equation}
	In general we have $
		-6\lambda^2 + 6\lambda - 1
		= \frac{c}{2}
	$; for $\lambda = 2$ this gives $c = -26$. 
	\qedfull

	\item \textbf{Free Fermion CFT:}
	\begin{equation}
		S = \int \dd[2]{z}
			\psi_i\,\pdbar\psi^i,\quad
		\psi^i = \psi_i^*,\quad
		\psi_i = \psi_i(z)
	\end{equation}

	\begin{enumerate}
	\item Mode expansion of such chiral fermion is given by:
	\begin{equation}
		\psi_i = \sum_{
			k\in\mbb{Z}+\frac{1}{2}
		} \frac{b_{ik}}{z^{k+\frac{1}{2}}},\quad
		b_{ik}
		= \frac{1}{2\pi i} \oint \dd{z}
			z^{k-\frac{1}{2}} \psi_i
	\end{equation}
	Canonical quantization is achieved by simply imposing anti-commutation relations; this is justified by mapping the system onto a cylinder, then $b_{ik}$'s indeed map to modes on the spatial circle\footnote{
		This can be proven rigorously by considering operator equations like in the $bc$ CFT problem. 
	}. The only non-zero commutators are:
	\begin{equation}
		\Bqty{b_{ik},b^{j\dagger}_q}
		= \delta_{k+q,0}\,\delta^j_i
	\end{equation}

	This gives the only non-zero 2-point functions:
	\begin{equation}
	\begin{aligned}
		\ave{
			\psi_i(z)\,\psi^j(w)
		}
		&= \sum_{k,q\in\mbb{Z}+\frac{1}{2}}
			\frac{1}{z^{k+\frac{1}{2}}}
			\frac{1}{w^{q+\frac{1}{2}}}
			\ave{b_{ik} b^{j\dagger}_q} \\
		&= \sum_{k,q\in\mbb{Z}+\frac{1}{2}}
			\frac{1}{z^{k+\frac{1}{2}}}
			\frac{1}{w^{q+\frac{1}{2}}}
			\mel{0}{
				\Bqty{b_{ik},b^{j\dagger}_q}
			}{0}
		= \frac{\delta^j_i}{z-w}
	\end{aligned}
	\end{equation}
	Note that $
		b^i_k\ket{0} = 0,\ %
		\forall\ k\ge\frac{1}{2}
	$.

	\item[(b)(c)] Combining two $\psi$ expansions gives the mode expansion of $
		J\id{_i^j}
		= \normorder{\psi_i(z)\,\psi^j(z)}
	$, namely:
	\begin{equation}
		J\id{_i^j}(z)
		= \sum_{k\in\mbb{Z}}
			\frac{(J\id{_i^j})_k}{z^{k+1}},\quad
		(J\id{_i^j})_k
		= \sum_{q\in\mbb{Z}+\frac{1}{2}}
			\normorder{
				b_{iq}\,b^{j\dagger}_{k-q}
			}
	\end{equation}
	It is in fact more convenient to obtain the $JJ$ OPE first, and then use it to find the  $[J_0,J_0]$ mode commutator\footnote{
		I would like to thank \textit{谷夏} for providing this hint. 
	}; note that $
		{
			{\psi}_i(z)\,{\psi}^j(w)
		}
	$ contraction gives $\frac{\delta^j_i}{z-w}$, we have:
	\begin{equation}
		J\id{_i^j}(z)\,J\id{_k^l}(0)
		\sim \frac{
			\delta_i^l\delta_k^j
		}{z^2} + \frac{
			\delta^j_k J\id{_i^l}(0)
			- \delta^l_i J\id{_k^j}(0)
		}{z},\\[1ex]
		\bqty{
			(J\id{_i^j})_0, (J\id{_k^l})_0
		}
		= \frac{1}{(2\pi i)^2}
			\oint_0 \dd{w}
			\oint_w \dd{z}
				J\id{_i^j}(z)\,J\id{_k^l}(w)
		= \delta_i^l\,(J\id{_k^j})_0
			- \delta_k^j\,(J\id{_i^l})_0
	\end{equation}

	\item[(d)] Similar to $bc$ CFT, we have:
	\begin{equation}
		T(z) = \frac{1}{2}\,\pqty{
			\normorder{\psi_i\,\pd\psi^i}
			- \normorder{\pd\psi_i\,\psi^i}
		},\quad
		T(z)\,T(w)
		\sim \frac{n/2}{(z-w)^4}
			+ \frac{2T(w)}{(z-w)^2}
			+ \frac{\pd T(w)}{z-w}
	\end{equation}
	With each (complex) field contributing $\frac{1}{2}\times 2$ central charge\footnote{
		In fact a complex (Dirac) fermion can be ``treated like'' (\textit{dual to}) a boson; this is \textit{bosonization}. 
	}.

	\item[(e)] For real fermions, there is an additional reality condition:
	\begin{equation}
		\psi^i = \psi^\star_i = \psi_i 
	\end{equation}
	The canonical quantization still holds without the extra adjoint, same as the 2-point function:
	\begin{equation}
		\ave{\psi_i(z)\,\psi_j(w)}
		= \frac{\delta_{ij}}{z - w}
	\end{equation}

	Similar holds for $J_{ij} = \normorder{\psi_i\psi_j}$ and its OPE, but we no longer need to distinguish upper/lower indices; we have:
	\begin{equation}
		J_{ij}(z)\,J_{kl}(0)
		\sim \frac{
			- \delta_{ik}\delta_{jl}
			+ \delta_{il}\delta_{jk}
		}{z^2} + \frac{
			- \delta_{ik} J_{jl}(0)
			+ \delta_{il} J_{jk}(0)
			+ \delta_{jk} J_{il}(0)
			- \delta_{jl} J_{ik}(0)
		}{z}\\[1ex]
		\bqty{
			(J_{ij})_0, (J_{kl})_0
		}
		= - \delta_{ik} (J_{jl})_0
			+ \delta_{il} (J_{jk})_0
			+ \delta_{jk} (J_{il})_0
			- \delta_{jl} (J_{ik})_0
	\end{equation}
	This is precisely the $\mfrak{o}(n)$ algebra. 
	\qedfull

	\end{enumerate}

	\end{enumerate}

\printbibliography[%
%	title = {参考文献} %
	,heading = bibintoc
]
\end{document}