### 2021-03-14 alpha/GravitationalEntropy/entropy.tex
Add note on gravitational entropy
... & update TeXplates for better cite style

% !TeX document-id = {b5392a94-51a3-49d1-9ba5-698bc09f9d35}
% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

\documentclass[a4paper
	,10pt
%	,twoside
]{article}

\input{../preamble.tex}
\input{../.modules/basics/biblatex.tex}

\title{Notes on Gravitational Entropy}
\addbibresource{entropy.bib}

%%% ID: sensitive, do NOT publish!
\InputIfFileExists{id.tex}{}{}

\begin{document}
\maketitle
\pagenumbering{arabic}
\thispagestyle{empty}

%\vspace*{-.5\baselineskip}

\setlength{\parskip}{.1\baselineskip}
\tableofcontents
\setlength{\parskip}{\parskipnorm}

%\addtocounter{section}{-1}
\section{Entanglement Entropy}
	Ref: \textcite{Apolo:2020qjm}.

	A boundary interval $\mcal{A} \Leftrightarrow \rho_\mcal{A} = e^{-\mcal{H}_\mcal{A}}$. $\mcal{H}_\mcal{A}$ is generically nonlocal, but in a highly symmetric background (e.g.~vaccuum) it might be geometrically realized and corresponds to a symmetry of the underlying quantum field theory.
	In this case there is a generalized Rindler transformation $f$ that maps the casual $\mcal{D} \to \{(\tau,\tilde{x})\}$ in some noncompact generalized Rindler spacetime, with $\pdd{\tau}$ an isometry of the spacetime. Then we have:
	\begin{equation}
		\zeta = 2\pi \pdd{\tau}
		= 2\pi a^i H_i,
	\quad
		\tau \sim \tau + 2\pi i
	\end{equation}
	$H_i$: symmetry generators. Vaccuum$|_\mcal{D}\to$ some thermal state with well-defined temperature $2\pi$.

\section{Entropy as Noether Charge}
	Ref: \textcite{Wald:1993nt,Lewkowycz:2013nqa,Faulkner:2013ana}.

\vspace{1.2\baselineskip}
\raggedright
\printbibliography[%
%	title = {参考文献} %
	,heading = bibintoc
]
\end{document}