### 2021-06-24 alpha/SUSY/susy.tex

% !TeX document-id = {b5392a94-51a3-49d1-9ba5-698bc09f9d35}
% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

\documentclass[a4paper
	,10pt
%	,twoside
]{article}

\input{preamble.tex}

\title{Understanding Supersymmetry}
\addbibresource{susy.bib}

%%% ID: sensitive, do NOT publish!
\InputIfFileExists{id.tex}{}{}

\makeatletter
\newcommand{\nobeginpar}{\@beginparpenalty=10000}
\makeatother

\begin{document}
\maketitle
\pagenumbering{arabic}
\thispagestyle{empty}

%\vspace*{-.5\baselineskip}

\setlength{\parskip}{.1\baselineskip}
\tableofcontents
\setlength{\parskip}{\parskipnorm}

\addtocounter{section}{-1}
\section{References}
\raggedright
	\begin{itemize}
	\item \cite{Atanasov:2017abc}~\fullcite{Atanasov:2017abc}
		\begin{itemize}
		\item \textbf{Appendix E} of: \par
		\cite{Belavin:1984vu}~\fullcite{Belavin:1984vu}.

		\item \textbf{Section \Romannum{4}} of: \par
		\cite{Kogut:1979wt}~\fullcite{Kogut:1979wt}.

		\item \textbf{Chapter 5} of:\par
		\cite{Fradkin:1991nr}~\fullcite{Fradkin:1991nr}.

		\end{itemize}
	\end{itemize}
\justifying

\section{Algebra \& Unitary Representations}

\vspace{1.2\baselineskip}
\pagebreak[4]
\raggedright
\printbibliography[%
%	title = {参考文献} %
	,heading = bibintoc
]
\end{document}