### 2023-06-20 alpha/GlueonAdS/glueon-beamer.tex
Tune glueon slides for physics audience

% !TeX document-id = {c68ea1ac-ebac-4541-a297-17005c6d2297}
% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TS-program = pdflatex
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

\documentclass[
	10pt
%	,draft
	,handout
	,noamsthm
]{beamer}

\input{preamble.tex}

\usepackage{tikz}
%\usepackage{snapshot}

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}
\author{%
	Wen-Xin Lai 
	\texstringonly{\,}\textkai{赖文昕}%
	\texstringonly{
		\textit{\small in collaboration with} \\
			Luis Apolo,
			Peng-Xiang Hao \textkai{郝鹏翔}
			and Wei Song \textkai{宋伟}\\[2ex]
		\sidenote{\arxiv{2303.04836}}%
%		\ submitted to \textit{JHEP}
	}%
}
\institute{\small%
	Yau Mathematical Sciences Center,\\
	Tsinghua\texstringonly{
%		\\\large\ccbyncsajp
	}%
}
\date{}

\newcommand{\veccol}[1]{\pqty{
	\begin{smallmatrix}
		#1
	\end{smallmatrix}
}}

\addbibresource{glueon-beamer.bib}
%\usepackage{cprotect}

\usepackage{cancel}
\newcommand{\slot}{{\,\bullet}}

\usepackage{xspace}
\newcommand{\TTbar}{\texorpdfstring{\ensuremath{T\bar{T}}}{TTbar}\xspace}

\begin{document}
{%%% TITLEPAGE
\setbeamercolor{title in head/foot}{
		use=palette quaternary
		,fg=palette quaternary.bg
	}
\logo{}
\begin{frame}
%	\titlegraphic{\vspace{2\baselineskip}}
	\titlepage
	\tikz [remember picture,overlay]
	\node at ([
		yshift=-4.8\baselineskip,
		xshift=.26\linewidth
	] current page.center) {
		\begin{minipage}{.5\textwidth}
		\centering
		\includegraphics[height=.33\textheight]{img/smbc-cantor-cropped.png}
		\\[.2ex]
		\footnotesize \url{https://smbc-comics.com/comic/cantor}
		\hspace{-1.5em}
		\end{minipage}
	};
\end{frame}
}%%% TITLEPAGE

\begin{frame}{Outline}
\raggedbottom
\large
\tableofcontents
\end{frame}

\section{\textbf{Warmup:} $\mrm{AdS}_{d+1}$/$\mrm{CFT}_{d}$ at $d = 2$}

\newcommand{\stateAdsCft}{
\begin{align*}
	\textrm{Strings on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{\textit{asympt.}~$\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at \textit{asympt.}~boundary}
\end{align*}
}

\newcommand{\citeMaldacena}{%
	\textcite{Maldacena:1997re} --- \scriptsize\citetitle{Maldacena:1997re}
}

\begin{frame}{AdS/CFT --- holography duality}{%
	\citeMaldacena%
}
\vspace{-2.5\baselineskip}
\stateAdsCft
\vspace{-1.8\baselineskip}
\begin{figure}[!h]
	\centering
	\includegraphics[width=.5\linewidth]{img/ads-cft.png}

	\vspace{-.3\baselineskip}
	\caption{$\mrm{AdS}_3/\mrm{CFT}_2$ cartoon}

	\vspace*{-.8\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{figure}
\vspace{-1.3\baselineskip}
\end{frame}

%\begin{equation}
%	R_{\mu\nu} - \frac{1}{2}\,R g_{\mu\nu}
%	+ \Lambda g_{\mu\nu}
%	= 8\pi G_N T_{\mu\nu}^{\textit{matter}},\quad
%	\Lambda < 0
%\label{eq:einstein}
%\end{equation}

\begin{frame}{AdS/CFT --- a model of quantum gravity}{%
	\citeMaldacena
}
\vspace{-2.4\baselineskip}
\stateAdsCft
\vspace{-1.6\baselineskip}
\pause
\begin{itemize}

\item A model of \textsc{quantum gravity}
\item ... originating from \textsc{string theory}
\item ... seemingly \textsc{Universal / ubiquitous}

\pause
\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}
	\item Our work: pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$ $(d=2)$.

\end{itemize}
\end{itemize}
\end{frame}

%%% LaTeX3e!
\NewDocumentCommand{\figAdsCft}{ O{.28} }{
\begin{column}{#1\textwidth}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/ads-cft-cleaner.png}

	\vspace{-.01\baselineskip}
	\scriptsize\textcite{AldegundePWSep22}\ \,
\end{figure}
\end{column}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a \textit{Poincar\'e cylinder}}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

	\vspace{-.5\baselineskip}
	boundary at $\rho \to 0$, $x^{i} \sim \varphi, t$, $i = 1,2$

\pause
	\item Surprisingly the expansion simply terminates at~$\rho\thinmspace g^{(4)}_{ij}$ for pure Einstein gravity! Moreover,
	\begin{equation}
		g^{(4)}_{ij} = \frac{1}{4} g^{(2)}_{ik} g^{(0)\thinmspace kl}_{\phantom{kl}} g^{(2)}_{lj}
	\end{equation}
	\textcite{Fefferman:2007rka,Banados:1998gg}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\newcommand{\eqBanados}{
	\begin{equation*}
	\hspace{-2em}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ \,
	\end{equation*}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Banados:1992wn}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric for $\rho > 0$, \\
	\textcite{Banados:1998gg}
	\eqBanados
	$g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
	where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\pause
	\item A consequence of Einstein equations \\
	in $(2+1)$ dimensions.\\
	\textit{Too many constraints, too little freedom!}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Different \textbf{geometries} in the ``bulk''
	\eqBanados
	specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: periodic functions, maps to different \textbf{states} in the boundary $\mrm{CFT}_2$ with stress energy $\ave{T(u)}_\psi$, $\ave{\bar{T}(v)}_\psi$

	\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities:\\
		\begin{itemize}
		\item spectrum,
			symmetry,\\
			partition functions,
			correlators, ...
		\end{itemize}
	\end{itemize}

\end{column}
\end{columns}
\end{frame}

\section{\textbf{Review:} \TTbar deformation and cutoff AdS holography} \label{se:cutoffholography}

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\newcommand{\citeTTbar}{%
	\textcite{Zamolodchikov:2004ce}, revived by \textcite{Smirnov:2016lqw} and \textcite{Cavaglia:2016oda} \textit{et al}
}

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

\item Define ``\,\TTbar\,'' as the flow of action:
\begin{equation}
	\partial_{\sidenote{\mu}} I = -{8 \pi} \int d^2x \, T\bar T =  -{\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)
	\label{TTbardef}
\end{equation}
For $\mrm{CFT}_2$,\ $T\bar{T} = T_{xx} T_{\bar{x}\bar{x}}$,\ where $x, \bar{x} = \varphi' \mp it'$.

\item The stress tensor $T_{ij}(\sidenote{\mu})$ on the right hand side flows with the deformation parametrized by $\sidenote{\mu}$.

\pause

\item This is thus a \textit{differential equation} for a flow, starting from some generic QFT specified by $I(\mu = 0)$. In this work we shall start from $\mrm{CFT}_2$, but note that \TTbar is \textit{irrelevant}: conformal symmetry will be broken.

\end{itemize}
\end{frame}

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

\item \textbf{``Solvable'':} the deformed spectrum of $\hat{H}(\mu)$ and $\hat{J}(\mu)$ on a cylinder of radius $R$ can be solved, 
and it's a simple function\\
 of the undeformed $E(0)$, $J(0)$:
\begin{equation}
\hspace{-2em}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0) % \label{ttbarspectrum}
\end{equation}
under simple (and reasonable) assumptions (e.g.~translation invariance et al).

\item \textbf{Variants:} suppose the theory has some ``components'' labeled by $w$,

\begin{itemize}
\item double-trace $\TTbar = (\sum_w T_w) (\sum_w \bar{T}_w)$ (this work)
\item single-trace $\TTbar' = \sum_w (T_w \bar{T}_w)$ (e.g.~\textit{\citeauthor{Cui:2023jrb}},~23)
\end{itemize}

\end{itemize}

\vspace{.5\baselineskip}
\end{frame}

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-1.\baselineskip}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,
yet it's still mysterious and difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

\pause
\item What if we start from a $\mrm{CFT}_2$, that enjoys a holographic dual to $\mrm{AdS}_3$...

\end{itemize}
\end{frame}

\newcommand{\figCutoffAds}[1][.3]{
\begin{column}{#1\textwidth}
\begin{figure}[!h]
	\hfill\includegraphics[width=\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\flushleft
	\ \scriptsize\textcite{AldegundePWSep22}
\end{figure}
\end{column}
\hspace{-2em}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-2em}
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item Question: what is the AdS dual of \TTbar deformation?\\

	\begin{itemize}
	\item look up \textit{double-trace deformations}\\
	in \textit{the dictionary}:\\
	\textcite{Heemskerk:2010hk}

\pause
	\item look at properties of \TTbar:\\
	signs of gravity and coarse-graining\\
	\textcite{Dubovsky:2012wk,Dubovsky:2013ira}
	\end{itemize}

	\pause
	\item Answer: AdS gravity within a finite Dirichlet wall

	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.7\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}

\pause

	\item Passed many non-trivial tests
	\item Related to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.75\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

	\begin{itemize}
		\item Dual to some ``solvable'' non-conformal theory
		\item A step towards quantum gravity in reality!
	\end{itemize}

\pause
	\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
	\end{equation}
	But \TTbar itself admits $\mu > 0$ with nice properties.\\
	\textit{What is the other side of the duality?}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\section{\textbf{Proposal:} Glue-on AdS holography} \label{se:glueonproposal}

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- analytic continuation}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric has only simple poles:
	\begin{equation}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ %\label{banados}
	\end{equation}
	thus it admits well-defined analytic continuation from $\rho > 0$ to $\rho \in \mbb{R}$.
	\item Continuation of the dictionary:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
	\ \quad
		\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
	\end{equation}
%	\vspace{-.5\baselineskip}

	\end{itemize}
\end{frame}

\newcommand{\figGlueon}{
\begin{figure}[!h]
	\centering
	\includegraphics[width=.8\linewidth]{img/diagram.pdf}
	\vspace{-.3\baselineskip}
	\caption{Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\ Top-down view of the ``Poincare cylinder''
\end{figure}
}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figGlueon
\end{column}
\begin{column}{.6\textwidth}
\vspace{-\baselineskip}
\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\vspace{-.5\baselineskip}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

	\item A prescription is required to ``renormalize'' the divergences
	\end{itemize}

\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Kraus:2018xrn} -- \citetitle{Kraus:2018xrn}
}
\begin{itemize}
\item Firstly, matching energy momentum \& the flow equations:
	\begin{equation}
		T_{ij}= \frac{\sigma_{\zeta_c}}{8\pi G} \Big( K_{ij} -  K \gamma_{ij} + \frac{1}{\ell |\zeta_c|}\gamma_{ij}\Big), \ \quad \sigma_{\zeta_c} = \frac{\abs{\zeta_c}}{\zeta_c} = \pm 1,
		\label{BYtensor}
	\end{equation}
	The field theory metric $\gamma_{ij} = \zeta_c h_{ij}$, where $h_{ij}$ is the induced metric.

	\pause

	\begin{itemize}
	\item $\gamma_{ij}$ is always positive-definite, while $h_{ij}$ becomes negative-definite for the glue-on region.\\
		This discrepancy is the origin of the absolute values.
	\end{itemize}

	\pause

\item \TTbar flow is recast geometrically as the $i,j$ components of the Einstein equations. \\
Note that the geometry can be decomposed such that:
	\begin{equation}
		ds^2 = \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu
	\end{equation}
\end{itemize}
\end{frame}

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- beyond the infinity}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{itemize}
\item The geometry is foliated by constant $\zeta$ surfaces ${\mathcal N}_\zeta$:
	\begin{equation}
	\begin{aligned}
		ds^2 
		&= \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg) \\
		&= \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu,\ \quad
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi}
	\end{aligned}
	\end{equation}
	The \TTbar deformed theory lives on $\mcal{N}_{\zeta_c}$ with metric $\gamma_{ij}$.

\pause
\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together\\
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}\qquad
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\centering
\vspace{-.5\baselineskip}
\begin{minipage}{.5\textwidth}
\centering
\includegraphics[width=.8\linewidth]{img/diagram.pdf}
\end{minipage}
\vspace{-.8\baselineskip}
\begin{equation}
\stateGlueon
\end{equation}
\end{frame}

\section{\textbf{Check:} \TTbar deformed charges \& partition functions} \label{se:charges}

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\pause
\item Demanding the extended geometry to be non-singular\\
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

\pause
\item We are careful to match the bulk \& boundary symmetries at $\mcal{N}_{\zeta_c}$

\item This reproduces $E(\mu)$, $J(\mu)$ with $\mu \in\mbb{R}$ in \eqref{ttbarspectrum}.

\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function \eqref{eq:partition} is approximated by the on-shell gravity action of the dominant saddle (method of steepest descent):
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

\pause
\item $I(\zeta_c)$ is roughly the volume of the geometry, which diverges at $\zeta \to 0^\pm$. We apply a natural extension of \textit{holographic renormalization} to remove the divergence.

\pause
\item $I = -\log \mcal{Z}$ satisfies the \TTbar flow \eqref{TTbardef}.\\
	This is non-trivial because we've packaged the quantum corrections of the boundary theory.
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\section{Summary \& future directions}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\vspace{-1.2\baselineskip}
\begin{equation}
\stateGlueon
\end{equation}
\vspace{-.3\baselineskip}
\begin{columns}
\begin{column}{.3\textwidth}
\centering
\includegraphics[width=\linewidth]{img/diagram.pdf}
\end{column}
\hspace{-1em}
\begin{column}{.7\textwidth}
\begin{itemize}
\item Holographic proposal for $T\bar T$-deformed CFTs with $\mu \in \mbb{R}$

\item As evidence, we show that the $T\bar T$ trace flow equation, the spectrum on the cylinder, and the partition function on the torus and the sphere, among other results, can all be reproduced from bulk calculations in glue-on AdS$_3$.

\pause
\item We hope to understand the entanglement structure of \TTbar deformed theories from bulk geometry.
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}

\centering

\huge
\sidenote{Thank you!}
\bigskip

\large
froms Wen-Xin Lai\ \textkai{赖文昕}

\textit{\normalsize in collaboration with}

	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}\\[2ex]
\sidenote{\arxiv{2303.04836}}
%\ submitted to \textit{JHEP}

\end{frame}

\begin{frame}[allowframebreaks]{References}
\renewcommand*{\bibfont}{\footnotesize}
\printbibliography %%% should not specify heading etc
\end{frame}

\end{document}